Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package \vskip15pt Supplementary Materials

### 2018-03-01 docs/MEE_submission/manuscript_MEE.tex
Updated supplement and MEE MS to try and cut down a bit

\subsection{Inter-observer variability in digitisations}

In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups ($n=10$ within groups). Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures. Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order. Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux). Digitisers varied in their level of experience, from people with experience of meta-analyses to those without any science background. We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).

As a measure of bias, we calculated the percentage differences from the true summary statistics as

\begin{equation}
\frac{\theta - \hat{\theta}}{\hat{\theta}}
\end{equation}

where $\theta$ is the estimate and $\hat{\theta}$ is the true value. The deviation from the true value of $r$ was not further standardised, as it is already on a standardised scale. We also took the absolute values of these standardised differences as a measure of precision. The resulting data was used to assess between- and within- user variability (i.e., the intra-class correlation coefficient - ICC). This was done using linear mixed effect models with user identity as a random effect using \pkg{lme4} \citep{bates2015} in \proglang{R}. Standardised mean, standard deviation and correlation coefficients were used as response variables in separate models. Sampling variance for ICC estimates was generated based on 1000 parametric bootstraps of the model and the significance was tested using likelihood ratio tests, using \pkg{rptR} \citep{stoffel2017}.

Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377). There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics. SD estimates from digitisations are clearly more prone to error than means or correlation coefficients. If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433). SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.

\subsection{Testing the accuracy of digitisations}
To test how accurate \pkg{metaDigitise} is at matching clicked points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}. This was done by one digitiser (J.L.P.), as there is no detectable between user variation. Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$).

>>>>

In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups. The same simulated datasets were given to 14 different digitisers to compare the inter-observer variability in digitisations. We also used simulated data to test the accuracy of digitisations with respect to known values (See SM for more details on how simulations were set up).

Across the plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377). There were was also little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics. SD estimates from digitisations are clearly more prone to error than means or correlation coefficients. If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433). SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.

\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly 
correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
